---
title: 'Creating PSC Market Briefing Charts: Agency Charts - QxQ - Civilian or Military Branch'
author: "Marielle Roth"
date: "August 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quarter by Quarter Agency Charts
### Civilian Agencies or Military Branches

This script will produce the quarter by quarter stacked bar chart for Civilian Agencies/Departments or Military Branches.  It won't work for DoD because the data pull is too big to download from BGOV.  See the **Agency Charts - QxQ - DoD** markdown file for creating the DoD quarter by quarter chart.

The files needed for these are saved in **X:/1 Marielle Folder/Data Sets/By Agency/Quarter by Quarter**.  If you need to pull an additional fiscal quarter of data, use BGOV:

+ Open the Contracts Intelliegence Tool
+ Click **+ More Fields**
+ Select **Transaction Signed Date**, then click **UPDATE**
+ Type the desired **Funding Agency** into the box and select
+ Scroll down to where it says **Transaction Signed Date** to specify for which quarter you want values
+ Click Download and select Download All
+ When the file finishes downloading open file, click the enable editing button, and click the tab at the bottom maked Data Sheet
+ Copy and paste the full dataset, except the first row (with the headings), into the correct dataset in **X:/1 Marielle Folder/Data Sets/By Agency/Quarter by Quarter** and rename it to reflect the correct dates
    + If the csv file has become too big for a single csv page, you may have to delete a previous year
    + If the full quarter is too big for BGOV to download, you may have the chunk the data and download it in smaller increments
    + If it's a new file, save using the same form as the other files in the folder, and save as a __CSV (Comma Delimited)__ file
    + You can double check the data after running the script by cross-referencing the values on the chart with those in BGOV (each year should be roughly the same)

After saving and renaming the file, open the script **Agency QxQ Charts - civ or mil branch** located in **X:/1 Marielle Folder/R Scripts/Useful PSC Scripts/QxQ Agency Charts**

First load the necessary packages and identify your location for saving your chart (set the working directory). Run the following code (*Highlight it* and click __Run__ in the top right corner to execute the highlighted code):

```{r, eval=FALSE}
library(tidyverse)
library(RColorBrewer)
options(scipen=999)

setwd("X:/1 Marielle Folder/Visualizations/Agency Charts/QbyQ charts") 
```

Then define the name of the Agency and Year in quotes, using the exact name and formatting used to save the file. See the following example (for a file saved as **DOJ FY17-19Q3.csv**):

```{r, eval=FALSE}
Agency <- "DOJ"
Year <- "FY17-19Q3"   
dis_year <- "FY17-19Q3" 

```
Note: the **Year** refers to the year range in the file name, exactly as written.  **dis_year** is the year span that you want in the chart and displayed in the title of the chart.  If you only wanted to show FY18 to FY19Q3 you would instead have:

```{r, eval=FALSE}
Agency <- "DOJ"
Year <- "FY17-19Q3"   
dis_year <- "FY18-19Q3" 
```
Make the necessary changes to the above to reflect the **Agency** and **Year** and **dis_year**, all in quotes, and run the code.

Next read the data into R by running the following:

```{r, eval=FALSE}
data <- read_csv(paste("X:/1 Marielle Folder/Data Sets/By Agency/Quarter by Quarter/", Agency," ", Year,".csv", sep = ""))
```

Process the data, by running the following code:

```{r, eval=FALSE}
data.organized <- data %>% 
  rename("transaction_value" = `Transaction Value`,
         "fiscal_quarter" = `Fiscal Quarter`) %>% 
  filter(`Fiscal Year` %in% c(2017:2019)) %>%                               
  select(transaction_value, fiscal_quarter) %>% 
  group_by(fiscal_quarter) %>%
  summarise(sum = sum(transaction_value)) %>% 
  separate(fiscal_quarter, into = c("FY","quarter"), sep = "Q") %>% 
  mutate(total_obligations = round((sum)/1000000000, digits=2)) %>%           
  group_by(FY) %>% 
  mutate(label_y = cumsum(total_obligations),
         prop = 100*total_obligations/sum(total_obligations))%>% 
  mutate(FYYear = paste("FY", FY, sep = "")) %>% 
  mutate(Q_quarter = paste("Q", quarter, sep =""))

```
Note: fill in the date range, without quarters, you want displayed in the chart in the line

    filter(`Fiscal Year` %in% c(2017:2019)) %>% 
    
This should reflect what was defined for dis_year.  If you want to show all available quarters for 2017, 2018, and 2019, enter the above.  If you want to show only 2018 and 2019, enter

    filter(`Fiscal Year` %in% c(2018:2019)) %>% 

Additionally, if the agency has smaller value contracts, the 1000000000 can be changed to 1000000 or 1000 make the plot easier to read. If you change it, keep it in mind when creating the plot below to change Billions to Millions or Thousands.  This would be changed in the line

     mutate(total_obligations = round((sum)/1000000000, digits=2)) %>%


Next, create the barplot. Run the following:
```{r, eval=FALSE}
plot <- 
  ggplot(data.organized, aes(x = FYYear, y = total_obligations, fill = factor(Q_quarter, levels = c("Q4","Q3", "Q2","Q1")))) +
  geom_bar(stat = "identity", color = "Black") +
  geom_text(aes(label = round(total_obligations, digits = 2), y = label_y), size = 4, vjust = 1.5, fontface = "bold")+
  geom_text(data = subset(data.organized
                          , FY != 2019
                          ), aes(label = sprintf('%.0f%%', prop), y = label_y), size = 4, vjust = 3, fontface = "bold")+
  stat_summary(fun.y = sum, aes(label = ..y.., group = FY),
               geom = "text", vjust = -.5, size = sum(4,1), fontface = "bold")+   
  scale_fill_manual(name = "Quarter", values = brewer.pal(9, "GnBu")[c(1,3,5,7)])+
  #facet_grid(~agency_comp, labeller = label_wrap_gen(20))+
  labs(y = "Contract Obligations (in) Billions",                                           
       title = paste(Agency, " Contract Obligations Comparison - ", dis_year, sep = ""),                 
       caption = "Data Source: Bloomberg Government")+
  theme(plot.title = element_text(hjust = 0.5, vjust = 3, size = 24, face = "bold"), 
        plot.subtitle = element_text(hjust = 0.5, size = 18, face = "bold"),
        plot.caption = element_text(size = 8, face = "italic"),
        axis.ticks.x = element_blank(),
        strip.text = element_text(face = "bold", size = 20), 
        axis.title.x = element_blank(),
        panel.spacing = unit(4, "lines"))

plot
```
 
 Two fields may need to be changed in the above.  

1.Whichever year has any incomplete year displayed, plug that into the following line, replacing the 2019 (if a year other than 2019):  

    , Year != 2019
If all years are complete, put a # before this line so that it reads

    #, Year != 2019

2. If you changed the dollar division in the last step, change **Billions** to **Millions** or **Thousands** in the line that reads 

    labs(y = "Contract Obligations (in) Billions",
    

The plot should appear in the window at the bottom right.  If it doesn't, make sure you ran all the lines above and that there are no typos in your changes.  to see it larger, push the **Zoom** button.

If the plot needs to be changed after viewing, make the necessary changes and rerun the script.  If satisfied with the plot, save it as a JPG ih the directory you specified above. To do so, run the following: 
 
 
```{r, eval=FALSE}
ggsave(filename = paste(Agency,
                        " Contract Obligations ", dis_year, " by quarter.jpg", sep = ""), plot,          
       width = 13, height = 6.5, units = "in")
```