---
title: 'Creating PSC Market Briefing Charts: Defense-Wide DPAP Chart'
author: "Marielle Roth"
date: "August 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Total Contract Spending, by DPAP and Products categories

Find the file called **DPAP (services and total) Data - Government Wide.csv** in the folder **X:/1 Marielle Folder/Data Sets/Government-Wide Data/csv/To Build/**

To update the dataset, use BGovMarkets pull down to obtain the values for the needed year for DoD, Non-DoD, and Government-Wide.  The Civilian (Non-DoD) value should be available, and stable, around October 15th.  Since DoD is 3 months behind, the DoD and Government-Wide numbers will be available and stable around January 15th.

The DPAP categories are 

* Construction Services
* Electronic & Communication Services
* Equipment Related Services
* Facility Related Services
* Knowledge Based Services
* Logistics Management Services
* Medical Services
* Research and Development
* Transportation Services

These cover all the services.  To get **Services** add each category together.  To get **Products** subtract **Services** from the total on BGOV (you can get this by just running all years and the **Total** will be the first line of numbers; to get Civilian total, subtract DoD from the total).  

The code is located in the R File **Government-Wide DPAP Chart - Government-Wide** within **X:\1 Marielle Folder\R Scripts\Useful PSC Scripts\Government-Wide Charts**

First load the necessary packages and identify your location for saving your chart (set the working directory) and read in your saved dataset. Run the following code (*Highlight it* and click __Run__ in the top right corner to execute the highlighted code):

```{r, eval=FALSE}
library(colorspace)
library(RColorBrewer)
library(tidyverse)
options(scipen=999)

setwd("X:/1 Marielle Folder/Visualizations/Government-Wide")

data <- read_csv("X:/1 Marielle Folder/Data Sets/Government-Wide Data/csv/To Build/DPAP (services and total) Data - Government Wide.csv")

data$`DPAP Category` <- factor(data$`DPAP Category`,
                               levels = c("Products", "Construction Services",
                                          "Electronic & Communication Services",
                                          "Equipment Related Services",
                                          "Facility Related Services",
                                          "Knowledge Based Services",
                                          "Logistics Management Services",
                                          "Medical Services",
                                          "Research and Development",
                                          "Transportation Services"),
                               ordered = is.ordered(data$`DPAP Category`))
```


####DoD Chart

Run the following code to process, plot and save the DoD chart.

```{r, eval=FALSE}

##Defense-Wide data####

##Process

DPAP_dod <- data %>%
  filter(`DPAP Category` != "Total") %>% 
  group_by(`Fiscal Year`) %>%
  arrange(desc(`DPAP Category`)) %>%
  #mutate(label_y = cumsum(`$ billions`))
  mutate(pors = ifelse(`DPAP Category`=="Products","Product","Service")) %>%
  group_by(`Fiscal Year`, pors) %>%
  mutate(`pors$` = sum(`$_billions_DoD`))


label_height_dod <- DPAP_dod %>%
  group_by(`Fiscal Year`, pors) %>%
  summarize(`pors$` = sum(`$_billions_DoD`)) %>%
  group_by(`Fiscal Year`) %>%
  arrange(desc(`pors`)) %>%
  mutate(label_y2 = cumsum(`pors$`)) %>%
  left_join(DPAP_dod, by = c("Fiscal Year", "pors", "pors$") )

##Plot

plot_dod <- ggplot(label_height_dod, aes(x = `Fiscal Year`, y = `$_billions_DoD`,
                                         fill = `DPAP Category`)) +
  geom_bar(stat = "identity") +
  geom_text(aes(x = `Fiscal Year`, label = round(`pors$`, digits = 2), y = label_y2), size = 4, vjust = 1.5, check_overlap = TRUE)+
  geom_bar(data = subset(label_height_dod, `DPAP Category` %in% c("Medical Services", "Products")), aes(x=`Fiscal Year`, y=`pors$`), colour = "black", fill = NA, stat = "identity"
  )+
  scale_fill_brewer(name = "Services/Products Contract Category", palette = "Set3") +
  labs(x="Fiscal Year", y = "Contract Obligations (in) Billions",
       title = paste("DoD", " Total Contract Spending", sep = ""))+
  theme(plot.title = element_text(hjust = 0.5, size = 24, face = "bold"), axis.ticks.x = element_blank())

##Save - Change to relevent year span when saving

ggsave(paste("Defense", " Total Contract Spending Service Product ", "FY09-FY18.jpg", sep = ""), plot_dod,
       width = 13, height = 6.5, units = "in")

```

When saving, make sure to change the **"FY09-FY18.jpg"** section to reflect the correct year.